---
title: "Singles in SHARE Data Prep"
author: "Alexander Schumacher"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(magrittr)
library(knitr)
library(haven)
```

# Data import SHARELIFE
I import data from the SHARELIFE relationship modules in waves 3 and 7. To account for the time between waves 3 and 7, I use the demographic and coverscreen data from waves 4 to 7.
```{r dta import}
library(haven)
rp_w3 <- read_dta("data/00_raw/w3/sharew3_rel8-0-0_rp.dta")

rp_w7 <- read_dta("data/00_raw/w7/sharew7_rel8-0-0_rp.dta")

# Load Waves 4 to 6 for info on marital status
dn4 <- read_dta("data/00_raw/w4/sharew4_rel8-0-0_dn.dta")
dn5 <- read_dta("data/00_raw/w5/sharew5_rel8-0-0_dn.dta")
dn6 <- read_dta("data/00_raw/w6/sharew6_rel8-0-0_dn.dta")
dn7 <- read_dta("data/00_raw/w7/sharew7_rel8-0-0_dn.dta")


# Load Waves 4 to 6 for info on partner in HH
cvr4 <- read_dta("data/00_raw/w4/sharew4_rel8-0-0_cv_r.dta")
cvr5 <- read_dta("data/00_raw/w5/sharew5_rel8-0-0_cv_r.dta")
cvr6 <- read_dta("data/00_raw/w6/sharew6_rel8-0-0_cv_r.dta")
cvr7 <- read_dta("data/00_raw/w7/sharew7_rel8-0-0_cv_r.dta")

table(rp_w3$sl_rp002_)
table(rp_w3$sl_rp002d_)
table(rp_w3$sl_rp016_)
```
I then construct variables determining whether a respondent has ever been married, cohabited or been in a long-term relationship.
```{r variable construction}

#Recode missings and change binary format

#W3
table(rp_w3$sl_rp002_, useNA="always")
rp_w3r <- rp_w3 %>% mutate_at(vars(sl_rp002_, sl_rp002d_, sl_rp016_), list(~case_match(., 1 ~ 1, 5 ~ 0)))
table(rp_w3r$sl_rp002_, useNA="always")
#W7
table(rp_w7$rp002_, useNA="always")
rp_w7r <- rp_w7 %>% mutate_at(vars(rp002_, rp002d_, rp016_), list(~case_match(., 1 ~ 1, 5 ~ 0))) 
table(rp_w7r$rp002_, useNA="always")

## Update W3 info with Waves 4 to 7

# Join info on marital status from DN modules
married_w3 <- left_join(rp_w3r, dn4, by="mergeid", suffix=c("", ".4"))
married_w3 <- left_join(married_w3, dn5, by="mergeid", suffix=c("", ".5"))
married_w3 <- left_join(married_w3, dn6, by="mergeid", suffix=c("", ".6"))
married_w3 <- left_join(married_w3, dn7, by="mergeid", suffix=c("", ".7"))


# Join info on cohabitation from coverscreen
married_w3 <- left_join(married_w3, cvr4, by="mergeid", suffix=c("", ".4"))
married_w3 <- left_join(married_w3, cvr5, by="mergeid", suffix=c("", ".5"))
married_w3 <- left_join(married_w3, cvr6, by="mergeid", suffix=c("", ".6"))
married_w3 <- left_join(married_w3, cvr7, by="mergeid", suffix=c("", ".7"))


# Select relevant variables
married_w3 <- select(married_w3, one_of(c("mergeid", "sl_rp002_", "sl_rp002d_",  "sl_rp016_", "dn014_", "dn014_.5", "dn014_.6", "dn014_.7", "dn040_", "dn040_.5", "dn040_.6", "dn040_.7", "partnerinhh", "partnerinhh.5", "partnerinhh.6", "partnerinhh.7")))

# Create index variables for cohabitation, marriage and LAT in any of the waves
married_w6 <- married_w3 %>%
  mutate(., newlywed = dn014_!=4 & !is.na(dn014_)) %>%
  mutate(., newlywed = dn014_.5!=4 & !is.na(dn014_.5) | newlywed==TRUE) %>%
  mutate(., newlywed = dn014_.6!=4 & !is.na(dn014_.6) | newlywed==TRUE) %>%
  mutate(., newlywed = dn014_.7!=4 & !is.na(dn014_.7) | newlywed==TRUE) %>%
  mutate(., roomies = partnerinhh==1) %>%
  mutate(., roomies = partnerinhh.5==1 | roomies==TRUE) %>%
  mutate(., roomies = partnerinhh.6==1 | roomies==TRUE) %>%
  mutate(., roomies = partnerinhh.7==1 | roomies==TRUE) %>%
  mutate(., grasswidow = dn040_==1) %>%
  mutate(., grasswidow = dn040_.5==1 | grasswidow==TRUE) %>%
  mutate(., grasswidow = dn040_.6==1 | grasswidow==TRUE) %>%
  mutate(., grasswidow = dn040_.7==1 | grasswidow==TRUE)


# Correct marriage variable with W4 to W7 data
table(married_w6$sl_rp002_, useNA="always")
married_w6$sl_rp002_[married_w6$newlywed==TRUE] <- 1
table(married_w6$sl_rp002_, useNA="always")

# Correct cohabitation variable with W4 to W7 data
table(married_w6$sl_rp002d_, useNA="always")
married_w6$sl_rp002d_[married_w6$roomies==TRUE] <- 1
table(married_w6$sl_rp002d_, useNA="always")

# Correct cohabitation variable with W4 to W7 data
table(married_w6$sl_rp016_, useNA="always")
married_w6$sl_rp016_[married_w6$grasswidow==TRUE] <- 1
table(married_w6$sl_rp016_, useNA="always")

#Join Waves
married_w6 <- select(married_w6, one_of(c("mergeid", "sl_rp002_", "sl_rp002d_", "sl_rp016_")))
married_w7 <- select(rp_w7r, one_of(c("mergeid", "rp002_", "rp002d_", "rp016_")))
married <- full_join(married_w6, married_w7, by="mergeid")


# Construct marriage variable from W7 info
married <- married %>%  mutate(evermarried=rp002_) 

# Fill missings with W3 info
married$evermarried[is.na(married$rp002_)] <-     married$sl_rp002_[is.na(married$rp002_)]

table(married$evermarried, useNA="always") 


# Construct cohabitation variable from W7 info
married <- married %>% mutate(cohabited=rp002d_)

# Fill missings with W3 info
married$cohabited[is.na(married$rp002d_)] <-     married$sl_rp002d_[is.na(married$rp002d_)]

# Use marriage info for cohabitation variable
married <- married %>% mutate(evercohab = if_else(evermarried==0, cohabited, evermarried, cohabited))


# Construct LAT variable from W7 info
married <- married %>% mutate(lat=rp016_)

# Fill missings with W3 info
married$lat[is.na(married$rp016_)] <-   married$sl_rp016_[is.na(married$rp016_)]

# Use cohabitation info for LAT variable
married <- married %>% mutate(everlat = if_else(evercohab==0, lat, evercohab, lat))
married <- married %>% select(mergeid, evermarried, evercohab, everlat)

table(married$evermarried, useNA="always") 
table(married$evercohab, useNA="always") 
table(married$everlat, useNA="always") 


```

Next, I add data on the Big 5 personality traits. Since the Big 5 have not been 
asked to SHARELIFE respondents in Wave 7, I supplement Wave 8 info on the Big 5
for these cases.
```{r personality traits}

# Load SHARE modules for Big 5
# W7
b5_w7 <- read_dta("data/00_raw/w7/sharew7_rel8-0-0_gv_big5.dta")
table(b5_w7$bfi10_extra, useNA="always") 

#W8
b5_w8 <- read_dta("data/00_raw/w8/sharew8_rel8-0-0_gv_big5.dta")
table(b5_w8$bfi10_extra, useNA="always") 

# Select relevant variables 
big5_w7 <- select(b5_w7, one_of(c("mergeid", "bfi10_extra", "bfi10_agree", "bfi10_consc", "bfi10_neuro", "bfi10_open")))
big5_w8 <- select(b5_w8, one_of(c("mergeid", "bfi10_extra", "bfi10_agree", "bfi10_consc", "bfi10_neuro", "bfi10_open")))


# Join waves 7 and 8
bigfive <- full_join(big5_w7, big5_w8, by="mergeid", suffix=c("", ".8"))

# Substitute missings in wave 7 by wave 8 data
bifi <- bigfive
for (i in 2:6) {
  j <- i+5
  bifi[is.na(bifi[,i]), i] <- bifi[is.na(bifi[,i]), j]
}

#Keep only relevant variables
bigfive <- bifi[,1:6]

table(bigfive[2], useNA = "always")

### item-level data -> for reliability
# Select relevant variables 
ac7 <- read_dta("data/00_raw/w7/sharew7_rel8-0-0_ac.dta")
ac8 <- read_dta("data/00_raw/w8/sharew8_rel8-0-0_ac.dta")

ac7 <- ac7 %>% select(mergeid, extra1 = ac701_, extra2 = ac706_, agree1 = ac702_, agree2 = ac707_, agree3 = ac711_,
                      consc1 = ac703_, consc2 = ac708_, neuro1 = ac704_, neuro2 = ac709_, openn1 = ac705_, openn2 = ac710_)
ac8 <- ac8 %>% select(mergeid, extra1 = ac701_, extra2 = ac706_, agree1 = ac702_, agree2 = ac707_, agree3 = ac711_,
                      consc1 = ac703_, consc2 = ac708_, neuro1 = ac704_, neuro2 = ac709_, openn1 = ac705_, openn2 = ac710_)

# Join waves 7 and 8
bigfive_items <- full_join(ac7, ac8, by="mergeid", suffix=c("", ".8"))

# Substitute missings in wave 7 by wave 8 data
bifi_items <- bigfive_items
for (i in 2:12) {
  j <- i+11
  bifi_items[is.na(bifi_items[,i]), i] <- bifi_items[is.na(bifi_items[,i]), j]
}

#Keep only relevant variables
bigfive_items <- bifi_items[,1:12]

table(bigfive_items[2], useNA = "always")
# recode b5 missings
bigfive_items <- bigfive_items %>% 
  mutate_at(vars(2:12),
                 list(~ ifelse( . %in% c(-8:-1), NA, .)))
```

I then add data on life satifaction, using both the CASP-12 index and the life
satisfaction question from the AC module. The CASP index seems to have a high 
number of missings (~4,500) and no data at all from Romania (another ~2000).
```{r life satisfaction}

# Load SHARE module gv health
gvhealth <- read_dta("data/00_raw/w7/sharew7_rel8-0-0_gv_health.dta")

# Keep CASP and mergeid columns
casp <- gvhealth %>% select(mergeid, casp)

# Load W7 AC module
ac <- read_dta("data/00_raw/w7/sharew7_rel8-0-0_ac.dta")

# Keep life satisfaction and mergeid columns
ac012 <- ac %>% select(mergeid, ac012_, #life satisfaction
                       casp1 = ac014_, casp2 = ac015_, casp3 = ac016_, casp4 = ac017_, casp5 = ac018_, casp6 = ac019_, # casp items
                       casp7 = ac020_, casp8 = ac021_, casp9 = ac022_, casp10 = ac023_, casp11 = ac024_, casp12 = ac025_)
# 1 [Often]   
# 2 [Sometimes]
# 3 [Rarely]  
# 4 [Never]   
table(ac012$ac012_, useNA = "always")

# Recode missings
ac012$ac012_[ac012$ac012_ < 0] <- NA
# recode casp missings
ac012 <- ac012 %>% 
  mutate_at(vars(3:14),
                 list(~ ifelse( . %in% c(-8:-1), NA, .)))

summary(casp) #many SHARE missings!! No data from Romania.
```

Next, I add info on age and gender of the respondents. For simplicity, age is 
calculated at the year 2017, regardless of whether the interview took place before 
or after the respondents' birthday.
``` {r age and gender}

#Load demographics data set 
dn <- read_dta("data/00_raw/w7/sharew7_rel8-0-0_dn.dta")

# Recode binary format
gender <- dn %>% 
  select(mergeid, dn042_) %>%
  mutate_at(vars(dn042_), list(~case_match(., 2 ~ 1, 1 ~ 0, -2 ~ NA, -1 ~ NA)))

gender$female <- gender$dn042_
gender <- gender[,c(1,3)]

# Calculate age (in 2017) from year of birth 
age <- dn %>% 
   mutate(age = 2017 - dn003_) %>%
  select(mergeid, dn003_, age, country)
  
summary(age$age)
```

I use the generated variables from the ISCED dataset for education.
``` {r education}

# Load ISCED generated variables data set
isced <- read_dta("data/00_raw/w7/sharew7_rel8-0-0_gv_isced.dta")
summary(isced$isced1997_r)

# Select relevant variables
educ <- isced %>%
  select(mergeid, isced1997_r)

# Recode missings
educ$isced1997_r[educ$isced1997_r < 0] <- NA

summary(educ$isced1997_r)

```

Income is taken from the imputations dataset. I solely use the first calculation 
for now. I am not sure whether this is best practice, but I will ask around and 
it should be easy to fix in case it is not. Note that the "thinc" variable has 
an implausible amount of 0 values. I would recommend to use thinc2, which does
not have this problem.
``` {r income}

# Load imputations data set
impu <- read_dta("data/00_raw/w7/sharew7_rel8-0-0_gv_imputations.dta")

# Use only the first of the imputation values
income <- impu[impu$implicat==1,] %>%
  select(mergeid, thinc2, thinc)
```

I then join all datasets, keeping only those cases that match. No IDs are lost 
in the process.
```{r merge datasets}
data1 <- full_join(gender, age, by ="mergeid")
summary(data1)
data2 <- inner_join(data1, married, by="mergeid")
data3 <- inner_join(data2, bigfive, by="mergeid")
data3i <- inner_join(data3, bigfive_items, by="mergeid")
data4 <- inner_join(data3i, ac012, by="mergeid")
data5 <- inner_join(data4, casp, by="mergeid")
data6 <- inner_join(data5, educ, by="mergeid")
data_prepped <- inner_join(data6, income, by="mergeid")
summary(data_prepped)
save(data_prepped, file = "data/01_cleaned/df_singles.rda")
```