---
title: "Singlehood SHARE: Analyses Overview"
author: "Michael Krämer"
toc: true
toc-depth: 5
toc-title: Contents
number-sections: true
execute:
  cache: true
format:
#  docx:
#    highlight-style: github
  html:
    link-external-icon: true
    link-external-newwindow: true
    code-block-bg: true
    code-block-border-left: "#31BAE9"
#comments:
#  hypothesis: true
editor: source
#bibliography: references-zotero.bib
#csl: apa.csl
---
<!--
You can find the preregistration [here](https://osf.io/xbua4).
-->

## Load packages

```{r}
#| echo: true
#| warning: false
#| code-fold: true
#| code-summary: "Show the code"
#| cache: false
library(renv)
library(tidyverse)
library(broom)
library(labelled)
library(psych)
library(GPArotation)
```

## Load data

```{r}
#| echo: true
#| warning: false
#| output: false
#| code-fold: true
#| code-summary: "Show the code"
#| cache: true
load("data/01_cleaned/df_singles.rda")
```

## Confirmatory results

### Hypothesis 1: Extraversion {#hyp1}

*Lifelong singles self-report lower levels of extraversion as compared to partnered individuals.* 

To examine this ...

#### Model

Computations:
```{r}
#| echo: true
#| warning: false
#| code-fold: true
#| code-summary: "Show the code"



```

Results:
```{r}
#| echo: true


```

::: {.callout-note appearance="minimal"}
interpretation
:::

### Hypothesis 2: Neuroticism {#hyp2}

*Lifelong singles self-report higher levels of neuroticism as compared to partnered individuals.* 

To examine this ...

#### Model

Computations:
```{r}
#| echo: true
#| warning: false
#| code-fold: true
#| code-summary: "Show the code"



```

Results:
```{r}
#| echo: true


```

::: {.callout-note appearance="minimal"}
interpretation
:::

### Hypothesis 3: Conscientiousness {#hyp3}

*Lifelong singles self-report lower levels of conscientiousness as compared to partnered individuals.* 

To examine this ...

#### Model

Computations:
```{r}
#| echo: true
#| warning: false
#| code-fold: true
#| code-summary: "Show the code"



```

Results:
```{r}
#| echo: true


```

::: {.callout-note appearance="minimal"}
interpretation
:::

### Hypothesis 4: Life Satisfaction {#hyp4}

*Lifelong singles self-report lower levels of life satisfaction as compared to partnered individuals.* 

To examine this ...

#### Model

Computations:
```{r}
#| echo: true
#| warning: false
#| code-fold: true
#| code-summary: "Show the code"



```

Results:
```{r}
#| echo: true


```

::: {.callout-note appearance="minimal"}
interpretation
:::