---
title: "Singlehood SHARE: Analyses Overview"
author: "Michael Krämer"
toc: true
toc-depth: 5
toc-title: Contents
number-sections: true
execute:
  cache: true
format:
#  docx:
#    highlight-style: github
  html:
    link-external-icon: true
    link-external-newwindow: true
    code-block-bg: true
    code-block-border-left: "#31BAE9"
#comments:
#  hypothesis: true
editor: source
#bibliography: references-zotero.bib
#csl: apa.csl
---
<!--
You can find the preregistration [here](https://osf.io/xbua4).
-->

## Load packages

```{r}
#| echo: true
#| warning: false
#| code-fold: true
#| code-summary: "Show the code"
#| cache: false
library(renv)
library(tidyverse)
library(broom)
library(labelled)
library(psych)
library(quest) # for POMP scores
library(GPArotation)
library(specr)
library(lme4)
library(lmerTest)
library(furrr)
library(knitr)
```

## Load cleaned data

```{r}
#| echo: true
#| warning: false
#| output: false
#| code-fold: true
#| code-summary: "Show the code"
#| cache: true
base::load("data/01_cleaned/df_singles.rda")
df_singles <- data_prepped
```

## More data cleaning

```{r}
#| echo: true
#| warning: false
#| output: false
#| code-fold: true
#| code-summary: "Show the code"
#| cache: true
df_singles <- df_singles %>% group_by(mergeid) %>% mutate(pid = cur_group_id()) %>% ungroup() %>% 
  mutate(age_c = age - mean(age, na.rm=T))

# rename for easier to interpret var names
df_singles <- df_singles %>% 
  rename(extraversion = bfi10_extra, agreeableness = bfi10_agree, conscientiousness = bfi10_consc, 
         neuroticism = bfi10_neuro, openness = bfi10_open,
         ever_partnered = everlat, ever_cohabitating = evercohab, ever_married = evermarried, 
         life_satisfaction = ac012_, quality_of_life = casp, income = thinc2)
# code education
df_singles <- df_singles %>% 
  mutate(education = case_when(
    isced1997_r == 1 ~ "ISCED_1",
    isced1997_r == 2 ~ "ISCED_2",
    isced1997_r == 3 ~ "ISCED_3",
    isced1997_r == 4 ~ "ISCED_4",
    isced1997_r == 5 ~ "ISCED_5",
    isced1997_r == 6 ~ "ISCED_6",
    isced1997_r %in% c(0, 95, 96) ~ "other/none",
    .default = NA_character_
  )) %>% # also dummy coding (1 = reference category)
  mutate(isced_2 = ifelse(!is.na(education), ifelse(education=="ISCED_2", 1, 0), NA),
         isced_3 = ifelse(!is.na(education), ifelse(education=="ISCED_3", 1, 0), NA),
         isced_4 = ifelse(!is.na(education), ifelse(education=="ISCED_4", 1, 0), NA),
         isced_5 = ifelse(!is.na(education), ifelse(education=="ISCED_5", 1, 0), NA),
         isced_6 = ifelse(!is.na(education), ifelse(education=="ISCED_6", 1, 0), NA),
         isced_other = ifelse(!is.na(education), ifelse(education=="other/none", 1, 0), NA)) %>% 
  mutate(income_z = scale(income)) # z-standardize income

# POMP scores for well-being (because scale is very different across the two!)
df_singles <- df_singles %>% 
  mutate(life_satisfaction_p = quest::pomp(life_satisfaction, mini=0, maxi=10, relative = FALSE, unit = 1),
         quality_of_life_p = quest::pomp(quality_of_life, mini=12, maxi=48, relative = FALSE, unit = 1))

# switch 0/1 dummy coding for predictors
df_singles <- df_singles %>% 
  mutate(never_partnered = if_else(ever_partnered==1, 0, ifelse(ever_partnered==0, 1, NA)),
         never_cohabitating = if_else(ever_cohabitating==1, 0, ifelse(ever_cohabitating==0, 1, NA)),
         never_married = if_else(ever_married==1, 0, ifelse(ever_married==0, 1, NA)))
```

## Merge country-level data

```{r}
#| echo: true
#| warning: false
#| output: false
#| code-fold: true
#| code-summary: "Show the code"
#| cache: true
df_singles <- df_singles %>% group_by(mergeid) %>% mutate(pid = cur_group_id()) %>% ungroup() %>% 
  mutate(country_str = as.character(haven::as_factor(country)))
df_singles$country_str <- sub(" ", "_", df_singles$country_str) # for Czech Republic

df_country_level <- read_csv2("country_data.csv") %>% rename(country_str = Country)

df_singles <- df_singles %>% left_join(df_country_level)

df_singles <- df_singles %>% 
  rename(religiosity = Religiosity, gender_ratio = `Gender_ratio(maleper100female)`, 
         singles_proportion = Total_Singlehood_Ratio) %>% 
  mutate(religiosity_z = scale(religiosity), # z-standardize 
         gender_ratio_z = scale(gender_ratio),
         singles_proportion_z = scale(singles_proportion))

# labelled variable types sometimes cause problems donwstream -> convert to numeric
df_singles <- df_singles %>% mutate_at(c("extraversion", "agreeableness", "conscientiousness", "neuroticism", "openness",
                                         "life_satisfaction_p", "quality_of_life_p"), as.numeric)
```

## Confirmatory results

### Specification curve analyses (SCA) {#sca}

Run SCA:
```{r}
#| echo: true
#| warning: false
#| output: false
#| code-fold: true
#| code-summary: "Show the code"
#| cache: true

specifications_b5 <- crossing(
  vec_outcome = c("extraversion", "agreeableness", "conscientiousness", "neuroticism", "openness"), 
  vec_predictor = c("never_partnered", "never_cohabitating", "never_married"),
  vec_controls = c("none", "age", "gender", "income", "education", "religiosity", "gender_ratio", "singles_proportion", "all")) %>% 
  mutate(specnr = row_number()) %>% 
  mutate(spec_name = paste0("outcome=", vec_outcome, ", predictor=", vec_predictor, 
                            ", controls=", vec_controls))

# custom function
# random intercept model:
lmer_ri_1 <- function(formula, data,...) {
  require(lme4)
  require(broom.mixed)
  formula <- paste(formula, "+ (1|country_str)")
  lme4::lmer(formula, data)
}
lmer_ri_2 <- function(formula, data,...) {
  require(lmerTest)
  require(broom.mixed)
  formula <- paste(formula, "+ (1|country_str)")
  lmerTest::lmer(formula, data)
}

# Including random slopes (only country as grouping variable)
lmer_rs_1 <- function(formula, data,...) {
  require(lme4)
  require(broom.mixed)
  slopevars <- unlist(strsplit(formula, " ~ "))[2]
  formula <- paste0(formula, "+ (1 + ", slopevars, "|country_str)" )
  lme4::lmer(formula, data, control=lmerControl(optimizer="bobyqa",
                                                optCtrl=list(maxfun=2e5)))
}
lmer_rs_2 <- function(formula, data,...) {
  require(lmerTest)
  require(broom.mixed)
  slopevars <- unlist(strsplit(formula, " ~ "))[2]
  formula <- paste0(formula, "+ (1 + ", slopevars, "|country_str)" )
  lmerTest::lmer(formula, data, control=lmerControl(optimizer="bobyqa",
                                                optCtrl=list(maxfun=2e5)))
}

# personality
# Setup Specifications
specs_b5 <- specr::setup(data = df_singles, 
                  y = c("extraversion", "agreeableness", "conscientiousness", "neuroticism", "openness"), 
                  x = c("never_partnered", "never_cohabitating", "never_married"), 
                  controls = c("age_c", "female", "income_z", "education", 
                               "religiosity_z", "gender_ratio_z", "singles_proportion_z"),
                  simplify = T, # not all combinations of covars
                  model = c("lmer_ri_1"))

# identical, but with p-values pls
specs_b5_p <- specr::setup(data = df_singles, 
                  y = c("extraversion", "agreeableness", "conscientiousness", "neuroticism", "openness"), 
                  x = c("never_partnered", "never_cohabitating", "never_married"), 
                  controls = c("age_c", "female", "income_z", "education", 
                               "religiosity_z", "gender_ratio_z", "singles_proportion_z"),
                  simplify = T, # not all combinations of covars
                  model = c("lmer_ri_2"))

# Run Specification Curve Analysis
#opts <- furrr_options( # for parallelization (using multiple cores)
#  globals = list(lmer_ri_1 = lmer_ri_1)
#)
#plan(strategy = multisession, workers = 6)
sca_b5 <- specr(specs_b5) 
                #.options = opts,   # Pass opts to specr
                #.progress = TRUE)
sca_b5_p <- specr(specs_b5_p) # run again (for p-values)

# well-being
# Setup Specifications
specs_wb <- specr::setup(data = df_singles, 
                         y = c("life_satisfaction_p", "quality_of_life_p"), 
                         x = c("never_partnered", "never_cohabitating", "never_married"), 
                         controls = c("age_c", "female", "income_z", "education", 
                                      "religiosity_z", "gender_ratio_z", "singles_proportion_z"),
                         simplify = T, # not all combinations of covars
                         model = c("lmer_ri_1"))

# identical, but with p-values pls
specs_wb_p <- specr::setup(data = df_singles, 
                         y = c("life_satisfaction_p", "quality_of_life_p"), 
                         x = c("never_partnered", "never_cohabitating", "never_married"), 
                         controls = c("age_c", "female", "income_z", "education", 
                                      "religiosity_z", "gender_ratio_z", "singles_proportion_z"),
                         simplify = T, # not all combinations of covars
                         model = c("lmer_ri_2"))

# Run Specification Curve Analysis
sca_wb <- specr(specs_wb) 
                #.options = opts,
                #.progress = TRUE)
sca_wb_p <- specr(specs_wb_p) # run again (for p-values)

#plan(sequential) # stop parallelization
```

SCA for personality traits:
```{r}
#| echo: true
#| fig-width: 9.5
#| fig-height: 7.5

# Plot Specification Curve
plot(sca_b5, 
     choices = c("x", "y", "controls"))
```

Summary of entire SCA:
```{r}
#| echo: true
summary(sca_b5)
```

SCA for well-being:
```{r}
#| echo: true
#| fig-width: 9.5
#| fig-height: 7.5

# Plot Specification Curve
plot(sca_wb, 
     choices = c("x", "y", "controls"))
```

Summary of entire SCA:
```{r}
#| echo: true
summary(sca_wb)
```

Important for interpretation: Unlike for personality, I used POMP (percentage of maximum possible) scores here to make the results comparable across the two well-being operationalizations.  

### Results summary

#### Hypothesis 1: Extraversion {#hyp1}

*Lifelong singles self-report lower levels of extraversion as compared to partnered individuals.* 

All model results for extraversion:
```{r}
#| echo: true
kable(sca_b5_p$data %>% filter(y=="extraversion") %>% 
        select(x, controls, estimate, conf.low, conf.high, std.error, statistic, p.value, fit_nobs), digits = 3)
```

Median effect size:
```{r}
#| echo: true
sca_b5$data %>% filter(y=="extraversion") %>% 
  group_by(x) %>% 
  summarise(median_est = median(estimate), median_conf.low = median(conf.low), median_conf.high = median(conf.high))
```

::: {.callout-note appearance="minimal"}
Positive effect for extraversion that does not really depend on the set of control variables. The effect is a bit stronger when looking at people who were ever in a committed relationship or cohabited with a partner compared to those who were ever married.
:::

### Hypothesis 2: Neuroticism {#hyp2}

*Lifelong singles self-report higher levels of neuroticism as compared to partnered individuals.* 

All model results for neuroticism:
```{r}
#| echo: true
kable(sca_b5_p$data %>% filter(y=="neuroticism") %>% 
        select(x, controls, estimate, conf.low, conf.high, std.error, statistic, p.value, fit_nobs), digits = 3)
```

Median effect size:
```{r}
#| echo: true
sca_b5$data %>% filter(y=="neuroticism") %>% 
  group_by(x) %>% 
  summarise(median_est = median(estimate), median_conf.low = median(conf.low), median_conf.high = median(conf.high))
```

::: {.callout-note appearance="minimal"}
No effect for neuroticism.
:::

### Hypothesis 3: Conscientiousness {#hyp3}

*Lifelong singles self-report lower levels of conscientiousness as compared to partnered individuals.* 

All model results for conscientiousness:
```{r}
#| echo: true
kable(sca_b5_p$data %>% filter(y=="conscientiousness") %>% 
        select(x, controls, estimate, conf.low, conf.high, std.error, statistic, p.value, fit_nobs), digits = 3)
```

Median effect size:
```{r}
#| echo: true
sca_b5$data %>% filter(y=="conscientiousness") %>% 
  group_by(x) %>% 
  summarise(median_est = median(estimate), median_conf.low = median(conf.low), median_conf.high = median(conf.high))
```

::: {.callout-note appearance="minimal"}
Quite consistent effect for conscientiousness that is a bit smaller than the one for extraversion but is relatively similar across all the different specifications.
:::

### Hypothesis 4: Well-being / Life Satisfaction {#hyp4}

*Lifelong singles self-report lower levels of life satisfaction as compared to partnered individuals.* 

Important for interpretation: Unlike for personality, I used POMP (percentage of maximum possible) scores here to make the results comparable across the two well-being operationalizations.  

All model results for well-being:
```{r}
#| echo: true
kable(sca_wb_p$data %>% filter(y %in% c("life_satisfaction_p", "quality_of_life_p")) %>% 
        select(y, x, controls, estimate, conf.low, conf.high, statistic, p.value, fit_nobs), digits = 3)
```

Median effect size:
```{r}
#| echo: true
sca_wb$data %>% filter(y %in% c("life_satisfaction_p", "quality_of_life_p")) %>% 
  group_by(y, x) %>% 
  summarise(median_est = median(estimate), median_conf.low = median(conf.low), median_conf.high = median(conf.high))
```

::: {.callout-note appearance="minimal"}
Substantial effect for both life satisfaction and quality of life. The effect is somewhat stronger for life satisfaction and, again (as for extraversion), for those who were ever in a committed relationship or cohabited with a partner compared to those who were ever married. Controlling for covariates (e.g., income, age) does make a substantial difference here.
:::

## Exploratory results

Differences in openness for experiences and agreeableness between lifelong singles and partnered individuals will be investigated in an exploratory manner.  

### Agreeableness

All model results:
```{r}
#| echo: true
kable(sca_b5_p$data %>% filter(y=="agreeableness") %>% 
        select(x, controls, estimate, conf.low, conf.high, std.error, statistic, p.value, fit_nobs), digits = 3)
```

Median effect size:
```{r}
#| echo: true
sca_b5$data %>% filter(y=="agreeableness") %>% 
  group_by(x) %>% 
  summarise(median_est = median(estimate), median_conf.low = median(conf.low), median_conf.high = median(conf.high))
```

::: {.callout-note appearance="minimal"}
Small effect but significant at *p* = .005. 
:::

### Openness to experiences

All model results:
```{r}
#| echo: true
kable(sca_b5_p$data %>% filter(y=="openness") %>% 
        select(x, controls, estimate, conf.low, conf.high, std.error, statistic, p.value, fit_nobs), digits = 3)
```

Median effect size:
```{r}
#| echo: true
sca_b5$data %>% filter(y=="openness") %>% 
  group_by(x) %>% 
  summarise(median_est = median(estimate), median_conf.low = median(conf.low), median_conf.high = median(conf.high))
```

::: {.callout-note appearance="minimal"}
Small effect that is not consistently significant for those never married.
:::

### Moderator analyses: age, income

*We will also investigate potential moderator variables for all hypothesized effects, as well as for openness for experiences and agreeableness.*  

In additional analyses, using multilevel regression models , we will investigate whether the relationship between singlehood and Big 5 personality traits or life satisfaction is moderated by participants' age or income/ SES.

Run models:
```{r}
#| echo: true
#| warning: false
#| output: false
#| code-fold: true
#| code-summary: "Show the code"
#| cache: true

# "+ (1|country_str)")
#   lmerTest::lmer(formula, data
```

### Moderator analyses: country-level variables

*Further, although there is some evidence for cross-country or cross-cultural differences in the association between being single and life satisfaction. Thus, we investigate this association in an exploratory manner as well.*

For the geographical analyses, we will investigate differences in life satisfaction and Big 5 personality traits among singles vs. non-singles regarding the country level variables (from the countries they are currently living in) religiosity, gender ratio and single ratio (level-2 variables). More precisely, we will investigate cross-level interaction effects of these level-2 variables with 'group'.


